Stats Project 2

### 2022-12-12 article.Rmd
Add RShiny App Files
updates to analysis 2 and selection models

---
title: "Stats Project 2"
author: Abhay Sharma | AbhaySh4r.github.io
output:
  pdf_document: default
  html_notebook: default
---

```{r include=FALSE}
library(tidyverse)
library(plotly)
library(car)
library(caret)
library(olsrr)
library(corrplot)
```

@@

```{=tex}
\setlength\parindent{18pt}
\section{Introduction}
```
For this project, we are provided a housing data set from the website Kaggle that contains a large collection of parameters that are likely to be correlated with the Sale Price of their respective houses. Within this data set we find 79 parameters, some of which must ultimately be used to build a predictive model to extrapolate housing prices given a certain set of these parameters.

@@

\subsection{Model Generation and Analysis}

To generate an initial model, we will look at a rudimentary approach to linear regression to see what our data does;

```{r echo = TRUE, results = 'hide'}
housing = read.csv(file = "train.csv")
a1_set = housing %>% select(SalePrice, Neighborhood, GrLivArea) %>% filter( Neighborhood == c("NAmes", "Edwards", "BrkSide"))
fit <- lm(SalePrice ~ GrLivArea + Neighborhood + GrLivArea*Neighborhood, data = a1_set)
summary(fit)
#plot(fit)
#plot(a1_set$SalePrice, a1_set$GrLivArea)

@@

```{r}

a1_set %>% ggplot(aes(x = (GrLivArea), y = SalePrice/1000, color = Neighborhood)) + geom_point() + stat_smooth(method = "lm")

a1_set %>% ggplot(aes(x = SalePrice/10000, fill = Neighborhood)) + geom_histogram(bins =20)

a1_set %>% ggplot(aes(x = GrLivArea, fill = Neighborhood)) + geom_histogram(bins =20)
```

@@

#put the plots here
plot(fit_transform)
```

We find that the log-log transform has significantly reduced our residuals and given us a much better fit for the linear regression model, represented by bolstering our R\^2 compared to the previous non-transformed model. We can see from the residual plots from the transformed fit that there are a few significantly high leverage/low residual points, and one out outlying point with a very high residual and leverage statistic. We can also verify from the adjusted $R^2$ and RMSE from each model that the transformed model fits the data much better than the a non-transformed model.

@@

For the second analysis in this project, we were tasked to generate 4 models using variations of parameters selected by Forward Selection, Backwards Elimination, Stepwise Selection, and a personalized approach. Using these paramters, we can generate Adjusted R\^2 values, CV Press statistics, and finally submit our model to the Kaggle website to obtain a "Kaggle Score" for each model.

```{r}
plot(t1fit)
```

To improve the quality of our model, we decided to remove outliers as analaysis shows marginal improvements to R\^2 and marginal deterioration to the Press Statistic. We notice the following outliers:

-   Excessively Expensive \> 700k SalePrice: ID #692 & \# 1183

-   Excessively Large \> 100k sq ft.: ID #314, #226, #256, #707

Next, we removed the following highly correlated variables to reduce multicollinearity: LotFrontage, MasVnrArea, GarageYrBlt, OverallQual, TotRmsAbvGrd, GarageCars, TotalBsmtSF, X1stFlrSF, X2ndFlrSF, FullBath. In the correlation matrix, we can see which plots are correlated. After the elimination of the variables mentioned, we see more fairly spare correlated variables.

```{r}
corrplot(cor(t1))
```

Stepwise Selection

Regression model:

Mean(SalePrice) = b0 + b1(OverallQual) + b2(OverallCond) + b3(YearBuilt) + b4(

## Model Results

| Model                | Adjusted R\^2 | RMSE      | CVPress  |
|----------------------|---------------|-----------|----------|
| Forward Selection    | 0.8333        | 0.1620974 | 43.77074 |
| Backward Elimination | 0.8333        | 0.1620974 | 43.77074 |
| Stepwise Selection   | 0.8333        | 0.1620974 | 43.77074 |

The rudimentary approach of numeric/reduced linearity set receives a Kaggle score of 0.20141. While is is interesting and good given the novel approach of the selection model, it's troubling to see that all 3 models returned the exact same Kaggle score, down to the last decimal. This may indicate that the models are adding/subtracting parameters to result in the exact same parameter list each time. This should be further investigated.

@@

```{r echo = TRUE, results = 'hide', fig.show= 'hide'}

@@

train.control1 = trainControl(method = "cv", number = 10)
model1 <- train(SalePrice ~ GrLivArea + Neighborhood + GrLivArea*Neighborhood, data = a1_set, method = "lm", trControl = train.control1)
CVPress(model1$model)# doesn't work atm

>>>>

---
title: "Stats Project 2"
author: Abhay Sharma | AbhaySh4r.github.io
output:
  pdf_document: default
  html_notebook: default
---
```{r include=FALSE}
library(tidyverse)
library(plotly)
library(car)
library(caret)
library(olsrr)
library(corrplot)
```

@@

\setlength\parindent{18pt}
\section{Introduction}

For this project, we are provided a housing data set from the website Kaggle that contains a large collection of parameters that are likely to be correlated with the Sale Price of their respective houses. Within this data set we find 79 parameters, some of which must ultimately be used to build a predictive model to extrapolate housing prices given a certain set of these parameters.

@@

\subsection{Model Generation and Analysis}
To generate an initial model, we will look at a rudimentary approach to linear regression to see what our data does;

```{r echo = TRUE, results = 'hide', warning=FALSE}
housing = read.csv(file = "train.csv")
numeric_housing = housing %>% select_if(is.numeric)
a1_set = housing %>% select(SalePrice, Neighborhood, GrLivArea) %>% filter( Neighborhood == c("NAmes", "Edwards", "BrkSide"))
fit <- lm(SalePrice ~ GrLivArea + Neighborhood + GrLivArea*Neighborhood, data = a1_set)
summary(fit)
#plot(fit)
#plot(a1_set$SalePrice, a1_set$GrLivArea)

@@

@@

We find that the log-log transform has significantly reduced our residuals and given us a much better fit for the linear regression model, represented by bolstering our R^2 compared to the previous non-transformed model. We can see from the residual plots from the transformed fit that there are a few significantly high leverage/low residual points, and one out outlying point with a very high residual and leverage statistic. We can also verify from the adjusted $R^2$ and RMSE from each model that the transformed model fits the data much better than the a non-transformed model.

@@

For the second analysis in this project, we have been tasked to build a predictive model that predicts the salary of houses provided the parameters of the dataset. The scope of this prediction is more general than the previous analysis question, focusing on the entire state of Iowa. Ultimately, the goal of these prediction is to compete against a known set on the website \textit{Kaggle} to receive a Kaggle score, demonstrating how accurate our prediction are.

\subsection{Model Selection and Generation}

Although we will be generating linear models for this analysis, the parameters for the model will be selected by a variety of selection methods, including, \textbf{forward selection, backwards elimination, and stepwise selection}. We will also be testing a set for which we select our own parameters to see how competitive we can get.

To generate the models, we must take an approach to limit the multicollinearity of the dataset, otherwise the selection models will not work. To do this, we can create a correlation plot of all the numeric values within the dataset, and visualize which values are highly correlated.

```{r}
transformed_numeric = numeric_housing
selected = subset(transformed_numeric, select = -c(LotFrontage, MasVnrArea, GarageYrBlt))
corrplot(cor(selected), tl.cex = 0.5)
```
From this plot, we can eyeball the selection of highly correlated values, including \textit{"OverallQual, TotRmsAbvGrd, GarageCars, TotalBsmtSF, X1stFlrSF, X2ndFlrSF, FullBath"}. By removing these variables from the dataset, we reduce the colinearity of the set and allow for a better feature selection process.

Running an initial backward selection model, we generate a series of residual plots that we can see below:

From the residual plots, we can see that due to the cloud-like random scattering of residuals, we've roughly explained the variations within the dataset with our model, this additionaly shows that there is no glaring breaks in variance. From the same plots, we can see that there is not sufficient evidence for any deviations from normality, seen in the QQ plots.We can additionalyly look at the VIFs for each model to numerically confirm that multicollinearity has been accounted for, with all numerical values falling below 3.

Results Table:

\begin{center}
\begin{tabular}{ |c|c|c|c| } 
 \hline
 Predictive Models & Adjusted R2 & CV PRESS & Kaggle Score \\ 
 \hline
 Forward & 0.8805 & 28.09 & 0.20061 \\
 \hline
 Backward & 0.8812 & 28.12 & 0.20061 \\
 \hline
 Stepwise & 0.8805 & 28.22 & 0.1814 \\
 \hline
 Custom & 
\end{tabular}
\end{center}

@@

```{r echo = TRUE, results = 'hide', fig.show= 'hide', warning=FALSE}

@@

train.control1 = trainControl(method = "cv", number = 10)
model1 <- train(SalePrice ~ GrLivArea + Neighborhood + GrLivArea*Neighborhood, data = a1_set, method = "lm", trControl = train.control1)
#CVPress(model1$model) # doesn't work atm

### 2022-12-12 rough_analysis.Rmd
Add RShiny App Files
updates to analysis 2 and selection models

```{r Reduced Linearity Set}
t1 = subset(transformed_numeric, select = -c(LotFrontage, MasVnrArea, GarageYrBlt, OverallQual, TotRmsAbvGrd, GarageCars, TotalBsmtSF, X1stFlrSF, X2ndFlrSF, FullBath )) 
#removal of largely correlated values to reduce multicollinearity (VIF /approx 1) and actually allow back-elimination to run,
#it wasn't running without a ton of errors without this.

@@

backt1 = ols_step_backward_p(t1fit, prem = 0.05, details = FALSE)
forwardt1 = ols_step_forward_p(t1fit, penter = 0.05, details = FALSE)               #the p-values for these should be fiddled with, 
stepwiset1 = ols_step_both_p(t1fit, pent = 0.05, prem =0.05, details = FALSE)

>>>>

```{r Reduced Linearity Set}
t1 = subset(transformed_numeric, select = -c(LotFrontage, MasVnrArea, GarageYrBlt, OverallQual, TotRmsAbvGrd, GarageCars, TotalBsmtSF, X1stFlrSF, X2ndFlrSF, FullBath ))

t1 = t1[-c(524,1299),] #removing outliers -524, 1299 for outlying lot area

#removal of largely correlated values to reduce multicollinearity (VIF /approx 1) and actually allow back-elimination to run,
#it wasn't running without a ton of errors without this.

@@

backt1 = ols_step_backward_p(t1fit, prem = 0.05, details = FALSE)
forwardt1 = ols_step_forward_p(t1fit, penter = 0.1, details = FALSE)               #the p-values for these should be fiddled with, 
stepwiset1 = ols_step_both_p(t1fit, pent = 0.05, prem =0.01, details = FALSE)
